---
title: "snowcam_treewell_data"
output: html_document
date: "2024-12-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse) # Package with dplyr, tibble, readr, and others to help clean coding
library(knitr) #Makes nice tables
library(kableExtra) #Makes even nicer tables
library(lubridate) #Makes working with dates easier
library(tidyr)
library(readxl)#Makes multiple simultaneous models easier
library(plotly)
library(rstatix)
library(ggpubr)
library(ggpmisc)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

Loading in Data

```{r}
snowcam_105E_transb_tw_data_edited <- read_csv("nsf/treewell_data/snowcam_treewell_data/105E_transb_tree_well_diams_edited.csv")

#this has original data with different tree diameters per date
snowcam_105E_transb_tw_data_full <- read_csv("nsf/treewell_data/snowcam_treewell_data/105E_transb_tree_well_diams.csv")

snowcam_LP_highmort_persb_tw_data_edited <- read_csv("nsf/treewell_data/snowcam_treewell_data/LP_highmort_persb_tree_well_diams_edited.csv")

#don't have this dataset done yet, so not ready to be loaded
snowcam_LP_lowmort_persb_tw_data_edited <- read_csv("nsf/treewell_data/snowcam_treewell_data/LP_lowmort_persb_tree_well_diams_edited.csv")
```

Data Wrangling

```{r}
#this is the version that has only one diameter per tree id
snowcam_transb_tw_edited <- snowcam_105E_transb_tw_data %>%
  mutate(date = mdy(date)) %>%
  mutate(well_diam_cm = as.numeric(well_diam_cm), tree_diam_cm = as.numeric(tree_diam_cm)) %>%  # Convert to numeric
  mutate(well_diam_cm = replace_na(well_diam_cm, 0)) %>%
  mutate(tree_diam_size = case_when(
    tree_diam_cm >= 20 ~ "large",
    tree_diam_cm >= 10 & tree_diam_cm < 20 ~ "medium",
    tree_diam_cm < 10 ~ "small",
    TRUE ~ NA_character_  # Handles cases where tree_diam_cm is NA
  ))
```

```{r}
#this is the version that is applying variable tree diameters per tree
snowcam_transb_tw_full <- snowcam_105E_transb_tw_data_full %>%
  mutate(date = mdy(date)) %>%
  mutate(well_diam_cm = as.numeric(well_diam_cm), tree_diam_cm = as.numeric(tree_diam_cm)) %>%  # Convert to numeric
  mutate(well_diam_cm = replace_na(well_diam_cm, 0)) %>%
  #making a column for tree diameter size - large = >20, medium = 10-20, small = <10
  mutate(tree_diam_size = case_when(
    tree_diam_cm >= 20 ~ "large",
    tree_diam_cm >= 10 & tree_diam_cm < 20 ~ "medium",
    tree_diam_cm < 10 ~ "small",
    TRUE ~ NA_character_  # Handles cases where tree_diam_cm is NA
  ))
```

### Data Exploration

Full Dataset with tree diams per date

```{r}
#making a dataframe that only has data for when there is snow at all trees
transb_snowon_treewell_data <- snowcam_transb_tw_full %>%
  filter(date <= as.Date("2024-04-23"))

#creating dataframes by tree_id
transb_snowon_treeid1_data <- snowcam_transb_tw_full %>%
  filter(tree_id == "tree_1") %>%
  filter(date <= as.Date("2024-04-23"))

transb_snowon_treeid2_data <- snowcam_transb_tw_full %>%
  filter(tree_id == "tree_2") %>%
  filter(date <= as.Date("2024-04-23"))

transb_snowon_treeid3_data <- snowcam_transb_tw_full %>%
  filter(tree_id == "tree_3") %>%
  filter(date <= as.Date("2024-04-23"))

transb_snowon_treeid4_data <- snowcam_transb_tw_full %>%
  filter(tree_id == "tree_4") %>%
  filter(date <= as.Date("2024-04-23"))
```

```{r}
lm_model <- lm(well_diam_cm ~ tree_diam_cm, data = transb_snowon_treewell_data)

r_squared <- summary(lm_model)$r.squared
```

```{r}
lm_model_treeid1 <- lm(well_diam_cm ~ tree_diam_cm, data = transb_snowon_treeid1_data)

r_squared_treeid1 <- summary(lm_model_treeid1)$r.squared
```

```{r}
lm_model_treeid2 <- lm(well_diam_cm ~ tree_diam_cm, data = transb_snowon_treeid2_data)

r_squared_treeid2 <- summary(lm_model_treeid2)$r.squared
```

```{r}
lm_model_treeid3 <- lm(well_diam_cm ~ tree_diam_cm, data = transb_snowon_treeid3_data)

r_squared_treeid3 <- summary(lm_model_treeid3)$r.squared
```

```{r}
lm_model_treeid4 <- lm(well_diam_cm ~ tree_diam_cm, data = transb_snowon_treeid4_data)

r_squared_treeid4 <- summary(lm_model_treeid4)$r.squared
```

```{r}
#scatterplot by tree diameter and well diameter

trans_tree_well_full_scatterplot <- ggplot(transb_snowon_treewell_data, aes(x = tree_diam_cm, y = well_diam_cm)) +
  geom_point(aes(color = tree_id), size = 3, alpha = 0.7) +  # Color by tree_id
  geom_smooth(method = "lm", color = "grey", se = TRUE) +  # Best-fit line with shaded CI
  annotate("text", x = 25, y = 10, label = paste("R² = ", round(r_squared, 2)), size = 5, color = "black") +# Add R² value
  #facet_wrap(~tree_diam_size) +
  labs(
    title = "Trans Burned Tree Diameter vs Well Diameter",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_minimal() +
  scale_color_viridis_d()  # Optionally, use a color palette for the tree_id

# Display the plot
trans_tree_well_full_scatterplot

```

Scatter plot by individual tree id - to see if R\^2 is better that way

```{r}
#treeid_1
trans_tree_well_treeid1_scatterplot <- ggplot(transb_snowon_treeid1_data) +
  geom_point(aes(x = tree_diam_cm, y = well_diam_cm, color = tree_id), color = "purple4", size = 3, alpha = 0.7) +  # Color by tree_id
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +  # Best-fit line with shaded CI
  annotate("text", x = 25, y = 30, label = paste("R² = ", round(r_squared_treeid1, 2)), size = 5, color = "black") +# Add R² value
  #facet_wrap(~tree_diam_size) +
  labs(
    title = "Tree id 1",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() +
  scale_x_continuous(
    limits = c(19, 27))

# Display the plot
trans_tree_well_treeid1_scatterplot
```

```{r}
#treeid_2
trans_tree_well_treeid2_scatterplot <- ggplot(transb_snowon_treeid2_data, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id), color = "gold", size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +  # Best-fit line with shaded CI
  annotate("text", x = 12, y = 15, label = paste("R² = ", round(r_squared_treeid2, 2)), size = 5, color = "black") +# Add R² value
  #facet_wrap(~tree_diam_size) +
  labs(
    title = "Tree id 1",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() +
  scale_x_continuous(
    limits = c(10, 15))

# Display the plot
trans_tree_well_treeid2_scatterplot
```

```{r}
trans_tree_well_treeid3_scatterplot <- ggplot(transb_snowon_treeid3_data, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id), color = "blue2", size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", color = "grey", se = TRUE) +  # Best-fit line with shaded CI
  annotate("text", x = 10, y = 8, label = paste("R² = ", round(r_squared_treeid3, 2)), size = 5, color = "black") +# Add R² value
  #facet_wrap(~tree_diam_size) +
  labs(
    title = "Tree id 3",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() +
  scale_x_continuous(limits = c(7, 11))

# Display the plot
trans_tree_well_treeid3_scatterplot
```

```{r}
trans_tree_well_treeid4_scatterplot <- ggplot(transb_snowon_treeid4_data, aes(x = tree_diam_cm, y = well_diam_cm)) +
  geom_point(aes(color = tree_id), color = "green4", size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", color = "grey", se = TRUE) +  # Best-fit line with shaded CI
  annotate("text", x = 20, y = 20, label = paste("R² = ", round(r_squared_treeid4, 2)), size = 5, color = "black") +  # Add R² value
  labs(
    title = "Tree id 4",  # Corrected the title to match the tree id
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() + 
  scale_x_continuous(limits = c(14, 22)) +
  scale_y_continuous() # Correctly added '+' before this line

# Display the plot
trans_tree_well_treeid4_scatterplot
```

```{r}
#patching all the tree_id plots together
transb_comb_bytreeid_scatterplot <- ggarrange(trans_tree_well_treeid1_scatterplot, trans_tree_well_treeid2_scatterplot, trans_tree_well_treeid3_scatterplot, trans_tree_well_treeid4_scatterplot, nrow = 2, ncol = 2)

transb_comb_bytreeid_scatterplot
```

Set tree diams by tree id - this doesn't make sense to do, because all the tree diams are the same per tree_id, which doesn't give me a scatterplot

```{r}
transb_snowon_treewell_edited_data <- snowcam_transb_tw_edited %>%
  filter(date <= as.Date("2024-04-23")
```

```{r}
trans_tree_well_edited_lineplot <- ggplot(transb_snowon_treewell_edited_data) +
  geom_line(aes(x = date, y = well_diam_cm, color = tree_id), linewidth = 0.8, alpha = 0.7) +  # Color by tree_id
  labs(
    title = "Trans Burned Treewell Developement",
    x = "Date",
    y = "Well Diameter (cm)"
  ) +
  theme_minimal() +
  scale_color_viridis_d()  # Optionally, use a color palette for the tree_id

# Display the plot
trans_tree_well_edited_lineplot

```

## V2 Data - processed by me

Download Data

```{r}
transb_snowcam_treewell_data = read_csv("nsf/treewell_data/snowcam_data/105E_transb_tree_well_diams_V2.csv")

persb_highmort_snowcam_treewell_data = read_csv("nsf/treewell_data/snowcam_data/LP_highmort_persb_tree_well_diams_V2.csv")

persb_lowmort_snowcam_treewell_data = read_csv("nsf/treewell_data/snowcam_data/LP_lowmort_persb_tree_well_diams_V2.csv")
```

```{r}
#snowcamera snow depth data
transb_snowcam_depth_data = read_csv("nsf/camera_snowdepths/2024_105E_transb_snowdepths.csv")

persb_highmort_snowcam_depth_data = read_csv("nsf/camera_snowdepths/2024_persb_lp_highmort_east_snowdepths.csv")

persb_lowmort_snowcam_depth_data = read_csv("nsf/camera_snowdepths/2024_persb_lp_lowmort_snowdepths.csv")
```

### Data Wrangling

```{r}
#edited snowcam data to match the timeline of the well data
transb_snowcam_depth_edited <- transb_snowcam_depth_data %>%
  filter(date >= "2024-04-01" & date <= "2024-05-09") %>%
  rename(burn_status = "mortality") %>%
  select(date, burn_status, zone, "maximum_depth(cm)", treewell_developed, conditions) %>%
  mutate(zone = "TSZ") %>%
  rename(max_depth_cm = "maximum_depth(cm)")

persb_highmort_snowcam_depth_edited <- persb_highmort_snowcam_depth_data %>%
  filter(date >= "2024-04-15" & date <= "2024-05-18") %>%
  select(date, mortality, burn_status, zone, max_depth_cm, treewell_developement, conditions) %>%
  mutate(zone = "PSZ") 

persb_lowmort_snowcam_depth_edited <- persb_lowmort_snowcam_depth_data %>%
  filter(date >= "2024-04-15" & date <= "2024-05-31") %>%
  select(date, mortality, burn_status, zone, max_depth_cm, conditions) %>%
  mutate(zone = "PSZ") 
```

```{r}
#snowcam treewell data edited
transb_snowcam_treewell_data_edited <- transb_snowcam_treewell_data %>%
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
  mutate(mortality = "mixed") %>%
  select(date, zone, burn_status, mortality, tree_id, tree_diam_cm, well_diam_cm)

persb_highmort_snowcam_treewell_data_edited <- persb_highmort_snowcam_treewell_data %>%
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
  select(date, zone, burn_status, mortality, tree_id, tree_diam_cm, well_diam_cm)

persb_lowmort_snowcam_treewell_data_edited <- persb_lowmort_snowcam_treewell_data %>%
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
  select(date, zone, burn_status, mortality, everything()) %>%
  select(date, zone, burn_status, mortality, tree_id, tree_diam_cm, well_diam_cm)
```

```{r}
#combined treewell dataframe 
full_snowcam_treewell_data <- rbind(persb_highmort_snowcam_treewell_data_edited, persb_lowmort_snowcam_treewell_data_edited, transb_snowcam_treewell_data_edited)
```

### Data Exploration

#### Line Plots

```{r}
#setting constants for tree_id colors and burn status linetypes
treeid_colors <- c(
  "tree_1" = "#1B7837",
  "tree_2" = "#D95F02",
  "tree_3" = "#2166AC",
  "tree_4" = "grey23"
)

burnstatus_linetype <- c("GB" = "solid", "DB" = "dashed")
```

```{r}
#data frame filtering out any well values of zero
transb_snowcam_treewell_data_nozeros <- transb_snowcam_treewell_data_edited %>%
  filter(well_diam_cm != 0) %>%
  mutate(burn_status = case_when(
    tree_id == "tree_4" ~ "DB",
    TRUE ~ burn_status))
```

```{r}
#plot by date vs treewell
##transitional
transb_tree_well_lineplot <- ggplot() + 
  geom_line(data = transb_snowcam_treewell_data_nozeros, aes(x = date, y = well_diam_cm, color = tree_id, linetype = burn_status), linewidth = 0.6) +
  # Bar plot for snow depth with custom legend
  geom_bar(data = transb_snowcam_depth_edited, 
           aes(x = date, y = max_depth_cm, fill = "daily max snow depth (cm)"),
           stat = "identity", color = NA, alpha = 0.3) + 
  labs(
    title = "Plot 1: Trans Burned",
    x = "Date",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() +
  scale_color_manual(values = treeid_colors) +
  scale_linetype_manual(values = burnstatus_linetype) +
  scale_fill_manual(name = "columns", values = c("daily max snow depth (cm)" = "grey32"))

# Display the plot
transb_tree_well_lineplot
```

```{r}
persb_highmort_snowcam_treewell_data_nozeros <- persb_highmort_snowcam_treewell_data_edited %>%
  filter(well_diam_cm != 0)
```

```{r}
##persb high mort
persb_highmort_tree_well_lineplot <- ggplot(persb_highmort_snowcam_treewell_data_nozeros) +
  geom_line(aes(x = date, y = well_diam_cm, color = tree_id, linetype = burn_status), linewidth = 0.6) +
   # Bar plot for snow depth with custom legend
  geom_bar(data = persb_highmort_snowcam_depth_edited, 
           aes(x = date, y = max_depth_cm, fill = "daily max snow depth (cm)"),
           stat = "identity", color = NA, alpha = 0.3) +
  labs(
    title = "Plot 2: Pers Burned High Mort",
    x = "Date",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() +
  scale_color_manual(values = treeid_colors) +
  scale_linetype_manual(values = burnstatus_linetype) + 
  scale_fill_manual(name = "columns", values = c("daily max snow depth (cm)" = "grey32"))

# Display the plot
persb_highmort_tree_well_lineplot
```

```{r}
persb_lowmort_snowcam_treewell_data_nozeros <- persb_lowmort_snowcam_treewell_data_edited %>%
  filter(well_diam_cm != 0)
```

```{r}
##persb low mort
persb_lowmort_tree_well_lineplot <- ggplot(persb_lowmort_snowcam_treewell_data_nozeros) +
  geom_line(aes(x = date, y = well_diam_cm, color = tree_id, linetype = burn_status), linewidth = 0.6) +
    # Bar plot for snow depth with custom legend
  geom_bar(data = persb_lowmort_snowcam_depth_edited, 
           aes(x = date, y = max_depth_cm, fill = "daily max snow depth (cm)"),
           stat = "identity", color = NA, alpha = 0.3) +
  labs(
    title = "Plot 3: Pers Burned Low Mort",
    x = "Date",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() + 
  scale_color_manual(values = treeid_colors) +
  scale_linetype_manual(values = burnstatus_linetype) + 
  scale_fill_manual(name = "columns", values = c("daily max snow depth (cm)" = "grey32"))

# Display the plot
persb_lowmort_tree_well_lineplot
```

```{r}
#combined dataframes
combined_treewell_lineplots <- ggarrange(transb_tree_well_lineplot, persb_highmort_tree_well_lineplot, persb_lowmort_tree_well_lineplot, nrow = 2, ncol = 2, common.legend = TRUE, legend = "right")

combined_treewell_lineplots
```

\*\* this data shows a pattern during the ablation periods of these three different sites, but those time periods are different at each site

-   Once the snow is mostly all melted, the well data is more sporadic, not as easy to see a signal between the date and well diameter

-   Correlation between snow depth and well depth and date for sure. As snow depth decreases, tree well depth increases.

\*\* Consider plotting net Radiation with these as well?

-   might make the most sense to ONLY show the peak ablation periods, and a bit of time before hand, since once the snow is mostly gone the signal gets weird.

#### Scatterplots

```{r}
#making dataframes that just have one tree ID per zone, so that we can do an R^2 assessment
transb_snowcam_treewell_data_id1 <- transb_snowcam_treewell_data_nozeros %>%
  filter(tree_id== "tree_1")

transb_snowcam_treewell_data_id2 <- transb_snowcam_treewell_data_nozeros %>%
  filter(tree_id== "tree_2")

transb_snowcam_treewell_data_id3 <- transb_snowcam_treewell_data_nozeros %>%
  filter(tree_id== "tree_3")

transb_snowcam_treewell_data_id4 <- transb_snowcam_treewell_data_nozeros %>%
  filter(tree_id== "tree_4")
```

```{r}
#treeid_1
trans_tree_well_id1_scatterplot_V2 <- ggplot(transb_snowcam_treewell_data_id1, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +  # Best-fit line with shaded CI
  #annotate("text", x = 12, y = 15, label = paste("R² = ", round(r_squared_treeid2, 2)), size = 5, color = "black") +# Add R² value
  #facet_wrap(~tree_diam_size) +
  labs(
    title = "Tree id 1",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw()
  #scale_x_continuous(
    #limits = c(10, 15))

# Display the plot
trans_tree_well_id1_scatterplot_V2
```

```{r}
#treeid_2
trans_tree_well_id2_scatterplot_V2 <- ggplot(transb_snowcam_treewell_data_id2, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +  # Best-fit line with shaded CI
  #annotate("text", x = 12, y = 15, label = paste("R² = ", round(r_squared_treeid2, 2)), size = 5, color = "black") +# Add R² value
  #facet_wrap(~tree_diam_size) +
  labs(
    title = "Tree id 2",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw()
  #scale_x_continuous(
    #limits = c(10, 15))

# Display the plot
trans_tree_well_id2_scatterplot_V2
```

```{r}
trans_tree_well_id3_scatterplot_V2 <- ggplot(transb_snowcam_treewell_data_id3, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +  # Best-fit line with shaded CI
  #annotate("text", x = 12, y = 15, label = paste("R² = ", round(r_squared_treeid2, 2)), size = 5, color = "black") +# Add R² value
  #facet_wrap(~tree_diam_size) +
  labs(
    title = "Tree id 3",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw()
  #scale_x_continuous(
    #limits = c(10, 15))

# Display the plot
trans_tree_well_id3_scatterplot_V2
```

```{r}
trans_tree_well_id4_scatterplot_V2 <- ggplot(transb_snowcam_treewell_data_id4, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +
  labs(
    title = "Tree id 4",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw()

# Display the plot
trans_tree_well_id4_scatterplot_V2
```

```{r}
#all points - trans burned 
trans_tree_well_alltrees_scatterplot <- ggplot(transb_snowcam_treewell_data_nozeros, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id, shape = burn_status), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +
  labs(
    title = "Plot 1: Trans Burned",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() + 
  scale_color_manual(values = treeid_colors) + 
  scale_shape_manual(values = c("DB" = 15, "GB" = 16))  # Star = 8, Dot = 16

trans_tree_well_alltrees_scatterplot
```

```{r}
#all points - pers burned high mort
persb_highmort_tree_well_alltrees_scatterplot <- ggplot(persb_highmort_snowcam_treewell_data_nozeros, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id, shape = burn_status), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +
  labs(
    title = "Plot 2: Pers Burned High Mort",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() + 
  scale_color_manual(values = treeid_colors) + 
  scale_shape_manual(values = c("DB" = 15, "GB" = 16))  # Star = 8, Dot = 16

persb_highmort_tree_well_alltrees_scatterplot
```

```{r}
#all points - pers burned low mort
persb_lowmort_tree_well_alltrees_scatterplot <- ggplot(persb_lowmort_snowcam_treewell_data_nozeros, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id, shape = burn_status), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +
  #facet_wrap(~date) +
  labs(
    title = "Plot 3: Pers Burned Low Mort",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() + 
  scale_color_manual(values = treeid_colors) + 
  scale_shape_manual(values = c("DB" = 15, "GB" = 16))  # Star = 8, Dot = 16

persb_lowmort_tree_well_alltrees_scatterplot
```

```{r}
comb_allzones_scatterplots <- ggarrange(trans_tree_well_alltrees_scatterplot, persb_highmort_tree_well_alltrees_scatterplot, persb_lowmort_tree_well_alltrees_scatterplot, nrow = 2, ncol = 2, common.legend = TRUE, legend = "right")

comb_allzones_scatterplots
```

Attempting to just do a short timeframe and see if there is a linear correlation then

```{r}
transb_april_1to5_data <- transb_snowcam_treewell_data_nozeros %>%
  filter(date >= "2024-04-01" & date <= "2024-04-05")
  
trans_tree_well_1to5_scatterplot <- ggplot(transb_april_1to5_data, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +
  labs(
    title = "Trans Burned: April 1-5th, 2024",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw()

# Display the plot
trans_tree_well_1to5_scatterplot
```

```{r}
transb_april_6to10_data <- transb_snowcam_treewell_data_nozeros %>%
  filter(date >= "2024-04-06" & date <= "2024-04-10")
  
trans_tree_well_6to10_scatterplot <- ggplot(transb_april_6to10_data, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +
  labs(
    title = "Trans Burned: April 6-10th, 2024",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw()

# Display the plot
trans_tree_well_6to10_scatterplot
```

```{r}
transb_april_11to15_data <- transb_snowcam_treewell_data_nozeros %>%
  filter(date >= "2024-04-11" & date <= "2024-04-15")

# Fit linear model
lm_model_transb_4_11to15 <- lm(well_diam_cm ~ tree_diam_cm, data = transb_april_11to15_data)

# Extract R-squared value
r_squared_transb_4_11to15 <- summary(lm_model)$r.squared
  
trans_tree_well_11to15_scatterplot <- ggplot(transb_april_11to15_data, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +
  labs(
    title = "Trans Burned: April 11-15th, 2024",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() + 
  # Annotate with R-squared value
  annotate("text", x = max(transb_april_11to15_data$tree_diam_cm) * 0.7, 
           y = max(transb_april_11to15_data$well_diam_cm) * 0.9, 
           label = paste("R² = ", round(r_squared, 2)), 
           size = 5, 
           color = "black")

# Display the plot
trans_tree_well_11to15_scatterplot
```

Peak Ablation Period Scatter Plots, facet wrapped by day

```{r}
#ablation period ONLY scatterplot

transb_peakablation_data <- transb_snowcam_treewell_data_nozeros %>%
  filter(date >= "2024-04-01" & date <= "2024-04-15")

trans_tree_well_peakablation_scatterplot_bydate <- ggplot(transb_peakablation_data, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id, shape = burn_status), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +
  facet_wrap(~date) +
  labs(
    title = "Trans Burned Peak Ablation: April 1-15th, 2024",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() +
  scale_color_manual(values = treeid_colors) +
  scale_shape_manual(values = c("DB" = 15, "GB" = 16))

trans_tree_well_peakablation_scatterplot_bydate
```

```{r}
#persb_highmort_peak ablation ONLY scatterplot
persb_highmort_peakablation_data <- persb_highmort_snowcam_treewell_data_nozeros %>%
  filter(date >= "2024-05-01" & date <= "2024-05-15")

persb_highmort_tree_well_peakablation_scatterplot_bydate <- ggplot(persb_highmort_peakablation_data, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id, shape = burn_status), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +
  facet_wrap(~date) +
  labs(
    title = "Pers Burned High Mort Peak Ablation: May 1-15th, 2024",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() +
  scale_color_manual(values = treeid_colors) +
  scale_shape_manual(values = c("DB" = 15, "GB" = 16))

persb_highmort_tree_well_peakablation_scatterplot_bydate
```

```{r}
#persb_highmort_peak ablation ONLY scatterplot
persb_lowmort_peakablation_data <- persb_lowmort_snowcam_treewell_data_nozeros %>%
  filter(date >= "2024-05-01" & date <= "2024-05-15")

persb_lowmort_tree_well_peakablation_scatterplot_bydate <- ggplot(persb_lowmort_peakablation_data, aes(x = tree_diam_cm, y = well_diam_cm)) +
 geom_point(aes(color = tree_id, shape = burn_status), size = 3, alpha = 0.7) +
  geom_smooth(method = "lm", aes(x = tree_diam_cm, y = well_diam_cm),  color = "grey", se = TRUE) +
  facet_wrap(~date) +
  labs(
    title = "Pers Burned Low Mort Peak Ablation: May 1-15th, 2024",
    x = "Tree Diameter (cm)",
    y = "Well Diameter (cm)"
  ) +
  theme_bw() +
  scale_color_manual(values = treeid_colors) +
  scale_shape_manual(values = c("DB" = 15, "GB" = 16))

persb_lowmort_tree_well_peakablation_scatterplot_bydate
```

\*\* splitting it by day shows that there is a correlation between diameter and well size, but that can only really be seen consistently on a day to day basis between different trees.

\*\* waiting on response from Stephanie before continuing to do this processing or adding it to my thesis!
